Read Me

### 2023-04-25 code/analysis_code/archieved/Race/p1_EDA_race.Rmd
modified outcome variable to MRR

Let's start by looking at the overall HIV-related death rate by race in the United States.
```{r}
race_df %>% 
  group_by(Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  mutate(percent_by_race = round(tot_deaths/sum(tot_deaths)*100, 1),
         ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population (>12 years old)",
                      "Death Rate (per 100,000)",
                      "Percent",
                      "Ratio"))
# using white persons as reference group
```

@@

```{r}
race_df %>% 
  ggplot() +
  geom_boxplot(aes(y = `Death Rate (estimated)`, x = Race)) + 
  labs(title = "HIV-Related Death Rate (calculated) by Race/ Ethnicity, United States, 2020", y = "Rate (per 100,000)", x = "Race/ Ethnicity") + theme(legend.position = "none")
```
The plots differ drastically for persons of multiple race. It is difficult to tell what might have caused this drastic change, but my inituition tells me there is something strange going on.

@@

ggplot(df, aes(y = outcome_DR, x = `Percent High School Education`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "HIV Death Rate vs. Percent High School Education") +
  facet_wrap(Race ~ . , scales = "free_y") +
  stat_cor(method="pearson")
ggsave(here("results/bi_ed.png"), units="in", width=10, height=7, dpi=300)
#"passed"

@@

ggplot(df, aes(y = outcome_DR, x = `Broadband of any type`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "HIV Death Rate vs. Broadband of any type") +
  facet_wrap(Race ~ . , scales = "free_y") + 
  stat_cor(method="pearson")
# did not "pass"
```

@@

ggplot(df, aes(y = outcome_DR, x = `Broadband of any type`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "HIV Death Rate vs. Broadband of any type") +
  facet_wrap(Race ~ . , scales = "free_y") + 
  stat_cor(method="pearson")
# did not "pass"

>>>>

Let's start by looking at the overall HIV-related death rate by race in the United States.
```{r}
race_df %>% 
  group_by(Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  mutate(percent_by_race = round(tot_deaths/sum(tot_deaths)*100, 1),
         ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population (>12 years old)",
                      "Death Rate (per 100,000)",
                      "Percent",
                      "Ratio"))
# using white persons as reference group

@@

```{r}
race_df %>% 
  ggplot() +
  geom_boxplot(aes(y = `Death Rate (estimated)`, x = Race)) + 
  labs(title = "HIV-Related Death Rate (calculated) by Race/ Ethnicity, United States, 2020", y = "Rate (per 100,000)", x = "Race/ Ethnicity") + theme(legend.position = "none")
```
The plots differ drastically for persons of multiple race. It is difficult to tell what might have caused this drastic change, but my intuition tells me there is something strange going on.

@@

ggplot(df, aes(y = outcome_DR, x = `Percent Less than High School Education`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "HIV Death Rate vs. Percent Less than High School Education") +
  facet_wrap(Race ~ . , scales = "free_y") +
  stat_cor(method="pearson")
ggsave(here("results/bi_ed.png"), units="in", width=10, height=7, dpi=300)
#"passed"

@@

ggplot(df, aes(y = outcome_DR, x = `Percent Without an Internet subscription`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "HIV Death Rate vs. Percent Without an Internet Subscription") +
  facet_wrap(Race ~ . , scales = "free_y") + 
  stat_cor(method="pearson")
# did not "pass"
```

@@

### 2023-04-25 code/analysis_code/archieved/Race/p2_PCA_race.Rmd
modified outcome variable to MRR

  # separating x variables and scaling
df_predictor <- df %>% 
    select(contains(c("State", "Race", "Percent", "Gini", "Median", "Broadband"))) %>% 
    #select(!contains(c("Virally", "Received", "Late"))) %>% 
    mutate(across(where(is.numeric), scale))

@@

#Scree plot
print(fviz_eig(pca, addlabels = TRUE))
ggsave(here("results/scree_plot.png"), units="in", width=10, height=7, dpi=300)
```
Since we originally had 11 variables, we should only look at components that explain more than 9.1% of variation. We can see from the scree plot that principal component 1 explained 63.8% of variation, and components 1 and 2 explained 84.7% of the variation, and components 1 - 3 explained 94.8%. Component 3 just  barely passed the threshold with 10.1%. All other principal components explains less that 6.2% of variation, so they will excluded.

>>>>

  # separating x variables and scaling
df_predictor <- df %>% 
    select(contains(c("State", "Race", "Percent", "Gini", "Median"))) %>% 
    select(!contains(c("Virally", "Received", "Late", "Percent High"))) %>% 
    mutate(across(where(is.numeric), scale))

@@

#Scree plot
print(fviz_eig(pca, addlabels = TRUE))
ggsave(here("results/scree_plot.png"), units="in", width=10, height=7, dpi=300)
```
Since we originally had 11 variables, we should only look at components that explain more than 9.1% of variation. We can see from the scree plot that principal component 1 explained 72.9% of variation, and components 1 and 2 explained 94.9% of the variation, and components 1 - 3 explained 94.8%. All other principal components explains less that 5.1% of variation, so they will excluded.

### 2023-04-25 code/analysis_code/archieved/Race/p3_model_race.Rmd
modified outcome variable to MRR

@@

Setting up functions for models

```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                 #       `Broadband of any type` + # estimate built 
                   #     `Percent Uninsured Under 65 Years Old`+ #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                     `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                    #     `Broadband of any type` + # estimate built 
                     #   `Percent Uninsured Under 65 Years Old`+ #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

@@

```{r}
  #poisson
pois_function <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                     `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                      #   `Broadband of any type` + # estimate built 
                     #   `Percent Uninsured Under 65 Years Old` + #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
  print(summary(pois_race))

@@

## function to fit GLM and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                     `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                     #    `Broadband of any type` + # estimate built 
                    #    `Percent Uninsured Under 65 Years Old` + #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
aic <- AIC(pois_race)
return(aic)
}
```

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                          `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                     #   `Broadband of any type` + # estimate built 
                      #  `Percent Uninsured Under 65 Years Old` + #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

@@

## function to fit GLM and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                          `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                      #   `Broadband of any type` + # estimate built 
                   #    `Percent Uninsured Under 65 Years Old` + #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

@@

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                         `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                       #  `Broadband of any type` + # estimate built 
                     #   `Percent Uninsured Under 65 Years Old`+ #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

@@

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                         `Percent Living in Poverty` +
                        `Percent High School Education` +  #estimate education access
                     #    `Broadband of any type` + # estimate built 
                    #    `Percent Uninsured Under 65 Years Old`+ #estimate access to healthcare
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

@@

```{r}
png(here("results/moran_black.png"))
#sink(here("results/spacy_summary_black.txt"))
spacy_function(black_df)
sink()
dev.off()
```

@@

```{r}
png(here("results/moran_hispanic.png"))
#sink(here("results/spacy_summary_hispanic.txt"))
spacy_function(hispanic_df)
sink()
dev.off()
```

```{r}
spacy_function(white_df)
```

```{r}
png(here("results/moran_white.png"))
#sink(here("results/spacy_summary_white.txt"))
spacy_function(white_df)
sink()
dev.off()
```

>>>>

```{r}
states <- data.frame(state.name)
states$region <- case_when(
  states$state.name %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania") ~ "Northeast",
  states$state.name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
  states$state.name %in% c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
  states$state.name %in% c("Arizona", "California", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
)

df <- full_join(df, states, by = join_by(State == state.name))
```

@@

Setting up functions for models
```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

@@

```{r}
  #poisson
pois_function <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
  print(summary(pois_race))

@@

## function to fit GLM and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                         `Gini Coefficient` + # estimate income inequality
                         `Percent Unemployed` +
                         `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
aic <- AIC(pois_race)
return(aic)
}
```

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                         `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

@@

## function to fit GLM and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

@@

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

@@

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

@@

```{r}
png(here("results/moran_black.png"))
sink(here("results/spacy_summary_black.txt"))
spacy_function(black_df)
sink()
dev.off()
```

@@

```{r}
png(here("results/moran_hispanic.png"))
sink(here("results/spacy_summary_hispanic.txt"))
spacy_function(hispanic_df)
sink()
dev.off()
```

### 2023-04-25 code/analysis_code/archieved/Race/p4_mapping.Rmd
modified outcome variable to MRR

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent High School Education`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent High School Education, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/ed_map.png"), units="in", width=10, height=7, dpi=300)

@@

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Living in Poverty` )) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Living in Poverty, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/poverty_map.png"), units="in", width=10, height=7, dpi=300)

@@

Let's map the observed/ expected ratio.
```{r}
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = OE_mortality_ratio)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) +
  facet_wrap(. ~ Race) +
  scale_fill_distiller(palette = "RdYlGn") + theme_void() + 
  theme(strip.text = element_text(size = 14),
    strip.background = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom") +
  labs(title = "Observed/ Expected HIV Mortality Ratio by Race/ Ethnicity in the United States, 2020", x = "", y = "", fill = "Ratio") +
  theme(plot.title= element_text(hjust = 0.5))

ggsave(here("results/roe_map.png"), units="in", width=10, height=4, dpi=300)
```
We can see that in much of the United States non-Hispanic Black persons had ratios consistently (much) greater than 1, suggesting a higher risk among this population.

@@

ggsave(here("results/fitted_map.png"), units="in", width=10, height=4, dpi=300)

```

>>>>

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Less than High School Education` )) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent of Adults with Less than High School Education, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/ed_map.png"), units="in", width=10, height=7, dpi=300)

@@

@@

@@

ggsave(here("results/fitted_map.png"), units="in", width=10, height=4, dpi=300)
```

### 2023-04-25 code/analysis_code/p4_mapping.Rmd
modified outcome variable to MRR

df_scaled <- readRDS(here("data/processed_data/df_fit.rds"))

df_raw <- readRDS(here("data/processed_data/race_new.rds"))

mod_aic <-  readRDS(here("data/processed_data/aic.rds"))
```

@@

#Best model  
best_mod_aic <- mod_aic %>%
  group_by(Race) %>%
  drop_na() %>% 
  filter(AIC == min(AIC))

df_scaled <- full_join(df_scaled, best_mod_aic, by = join_by("Race"))

##Pulling fitted values based on best fitting model
df_scaled <- df_scaled %>% 
  mutate(best_fitted = case_when(Model == "CAR" ~ car_fit,
                                 Model == "GLM" ~ mult_fit,
                                 Model ==  "Poisson" ~ pois_fit,
                                 Model == "Negative Binomial" ~ nb_fit))

@@

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Gini Coefficient`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Gini Coefficient, 2020", x = "", y = "", fill = "Gini coefficient (higher values represent greater income inequality)")  +
  theme(plot.title= element_text(hjust = 0.5))
ggsave(here("results/gini_map.png"), units="in", width=10, height=7, dpi=300)

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Less than High School Education` )) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent of Adults with Less than High School Education, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/ed_map.png"), units="in", width=10, height=7, dpi=300)

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Living with Severe Housing Cost Burden`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Living with Severe Housing Cost Burden, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/housing_map.png"), units="in", width=10, height=7, dpi=300)

ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Unemployed` )) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Unemployed, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/unemployed_map.png"), units="in", width=10, height=7, dpi=300)
```
The map for HIV-related death differs drastically across the United States by race/ethnicity. In most states, the highest HIV-related death were among Black person. This is especially true in the South region. Non-coastal western states consistently had lowest rates HIV-related deaths across the selected racial/ethnic groups, expect among multi-racial non-Hispanic persons. Non-Hispanic White persons were the only of the selected racial/ ethnic group that did not have a HIV-related death rate higher than ~10 per 100,000.

Next, let's map after the the Box_Cox transformation of the death rate
```{r}
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = BOX)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) +
  facet_wrap(. ~ Race) +
  scale_fill_distiller(palette = "RdYlGn") + theme_void() + 
  theme(strip.text = element_text(size = 14),
    strip.background = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom") +
  labs(title = "Death Rates (Box Cox transformed) by Race/ Ethnicity
in the United States, 2020", x = "", y = "", fill = "Ratio") +
  theme(plot.title= element_text(hjust = 0.5))
ggsave(here("results/roe_box_map.png"), units="in", width=10, height=4, dpi=300)
```

Next, let's map a probability map based on where we'd expect deaths if risk of HIV-related death was consistent across the US.
```{r}
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = Exp_DR)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) +
  facet_wrap(. ~ Race) +
  scale_fill_viridis_c() + theme_void() + 
  theme(strip.text = element_text(size = 14),
    strip.background = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom") +
  labs(title = "NULL - Expected HIV Mortality Rate by Race/ Ethnicity
       in the United States, 2020", x = "", y = "", fill = "Rate") +
  theme(plot.title= element_text(hjust = 0.5))
```
Well.. I probably need to go back and reconsider using making a lambda based on specific race/ ethnic. I think that at this point, however, it is still fair to consider mapping an observed/ expected ratio using the expected deaths calculate from the lambda for the entire population. This will allow us to see differences across the race/ ethnicity, rather than masking disparities in rates due higher expected cases among certain high-risk populations, which would thus make the ratio closer to the null.

@@

Next, let's make a map using the fitted values from the CAR model, the best performing model.
```{r}
##Smoothed map using fitted values from best models.
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = best_fitted)) +
  scale_fill_gradientn(colors = rev(brewer.pal(9, "RdYlGn"))) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) +
  facet_wrap(. ~ Race)  + 
  theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Smoothed Map (Fitted) of HIV-Related Death Rate by Race/ Ethnicity in the United States, 2020", x = "", y = "", fill = "") +
  theme(plot.title= element_text(hjust = 0.5))

>>>>

df_scaled <- readRDS(here("data/processed_data/df_fit_alt.rds"))
df_raw <- readRDS(here("data/processed_data/race_alt.rds"))
mod_aic <-  readRDS(here("data/processed_data/aic_alt.rds"))
```

@@

@@

```{r}
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`MRR`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio in the United States, 2020", x = "", y = "", fill = "Death rate per 100,000") +
  scale_fill_distiller(palette = "RdYlGn") +
  theme(plot.title= element_text(hjust = 0.5))

ggsave(here("results/mrr_map.png"), units="in", width=10, height=4, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Gini Coefficient`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Gini Coefficient, 2020", x = "", y = "", fill = "Gini coefficient (higher values represent greater income inequality)")  +
  theme(plot.title= element_text(hjust = 0.5))
ggsave(here("results/gini_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Living in Poverty`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Living in Poverty, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/poverty_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Uninsured Under 65 Years Old`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Uninsured Under 65 Years Old, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/uninsured_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Living with Severe Housing Cost Burden`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Living with Severe Housing Cost Burden 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/housing_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Living with Food Insecurity`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Living with Food Insecurity, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/foodinsecure_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Without an Internet subscription`)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Without an Internet Subscription, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/no_internet_map.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(df_sf_raw, aes(long, lat)) +
  geom_polygon(aes(group = group, fill =`Percent Unemployed` )) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)   + theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Percent Unemployed, 2020", x = "", y = "", fill = "%") +
  theme(plot.title= element_text(hjust = 0.5)) 
ggsave(here("results/unemployed_map.png"), units="in", width=10, height=7, dpi=300)
```
The map for HIV-related death differs drastically across the United States by race/ethnicity. In most states, the highest HIV-related death were among Black person. This is especially true in the South region. Non-coastal western states consistently had lowest rates HIV-related deaths across the selected racial/ethnic groups, expect among multi-racial non-Hispanic persons. Non-Hispanic White persons were the only of the selected racial/ ethnic group that did not have a HIV-related death rate higher than ~10 per 100,000.
`

Next, let's map a probability map based on where we'd expect deaths if risk of HIV-related death was consistent across the US.
```{r}
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = Exp_DR)) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) +
  facet_wrap(. ~ Race) +
  scale_fill_viridis_c() + theme_void() + 
  theme(strip.text = element_text(size = 14),
    strip.background = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom") +
  labs(title = "NULL - Expected HIV Mortality Rate by Race/ Ethnicity
       in the United States, 2020", x = "", y = "", fill = "Rate") +
  theme(plot.title= element_text(hjust = 0.5))
```

@@

```{r}
##Smoothed map using fitted values from best models.
ggplot(df_sf_scale, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = car_fit)) +
  scale_fill_gradientn(colors = rev(brewer.pal(9, "RdYlGn"))) +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)  + 
  theme_void() + 
  theme(strip.text = element_text(size = 14),
        legend.position = "bottom") +
  labs(title = "Smoothed Map (Fitted) of HIV-Related Death Rate by Race/ Ethnicity in the United States, 2020", x = "", y = "", fill = "") +
  theme(plot.title= element_text(hjust = 0.5))
